---
title: "R Notebook"
output:
  word_document: default
  pdf_document: default
  html_notebook: default
---

Created a Postgresql db and loaded the table with superstore store data and the script reads the table in R 

```{r}
library(DBI)
db <- 'practicedb'  
host_db <-'localhost'  
db_port <- '5432' 
db_user <- 'postgres'  
db_password <- 'root123'
con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password)

#check if the connection is established
dbListTables(con) 
 
#method returns an overview of the data stored in the database 
df <- dbGetQuery(con, 'SELECT * FROM superstore') 
head(df)
```

Checking the column names of the data, number of rows and columns

```{r}
names(df)
nrow(df)
ncol(df)
```

Removing NA's

```{r}
na.omit(df)

```
```{r}
tail(df$order_date)

```
Inner join in R using merge() function

```{r}
df1 <-data.frame(df$order_id,df$order_date,df$ship_date)
cols1<- c("order_id","order_date","ship_date")
colnames (df1)<-cols1

df2 <- data.frame (df$order_id,df$region,df$states,df$city)
cols2 <- c("order_id","region","states","city")
colnames(df2) <-cols2

inner <- merge(x=df1,y=df2, by="order_id")
inner

```
Left join in R using merge() function

```{r}
df3 <- data.frame(df$order_id,df$ship_mode,df$segment,df$category)
cols3 <-c("order_id","ship_mode","segment","category")
colnames(df3) <-cols3
  
df4 <- data.frame(df$order_id,df$sales,df$quantity,df$discount,df$profit)
cols4 <-c("order_id","sales","quantity","discount","profit")
colnames(df4) <-cols4

left <-merge(x=df3,y=df4,by="order_id",all.x=TRUE)
left
```
```{r}
df3 <- data.frame(df$order_id,df$ship_mode,df$segment,df$category)
cols3 <-c("order_id","ship_mode","segment","category")
colnames(df3) <-cols3
  
df4 <- data.frame(df$order_id,df$sales,df$quantity,df$discount,df$profit)
cols4 <-c("order_id","sales","quantity","discount","profit")
colnames(df4) <-cols4

right <-merge(x=df3,y=df4,by="order_id",all.y=TRUE)
right
```
Outer join in R using merge() function

```{r}
df3 <- data.frame(df$order_id,df$ship_mode,df$segment,df$category)
cols3 <-c("order_id","ship_mode","segment","category")
colnames(df3) <-cols3
  
df4 <- data.frame(df$order_id,df$sales,df$quantity,df$discount,df$profit)
cols4 <-c("order_id","sales","quantity","discount","profit")
colnames(df4) <-cols4

outer <-merge(x=df3,y=df4,by="order_id",all=TRUE)
outer
```

Pie chart 

concatenates labels and pct values and stores it in variable lbls

```{r}

# to only return sales that are positive
#df <- df[df$sales >= 0, ]
sales_by_category<- aggregate(df$sales, by=list(category=df$category), FUN=sum)


#sales_by_category<- data.frame(df$category,df$sales)
col<- c("category","sales")
colnames(sales_by_category) <- col
pct <- round(sales_by_category$sales/sum(sales_by_category$sales)*100)

labels <- df$category
lbls <-paste(labels,pct)
lbls <- paste(lbls,"%",sep="")
pie(table(sales_by_category$sales),labels=lbls,col=rainbow(3),main="sales_by_category_pie_chart")
legend("topright", sales_by_category$category, cex=0.8,fill=rainbow(length(sales_by_category$sales)))

```
The package plotrix has a function called pie3D() that is used for plotting a 3D Pie chart

explode is a parameter used to increase or decrease the gap of split in slices

```{r}
#install.packages("plotrix")
library(plotrix)
pie3D(sales_by_category$sales,labels=lbls,explode=0.05,main="sales_by_category")
```
bar chart

```{r}

#df_bar <-head(df_bar,n=10)
#df_bar %>% top_n(10)
df_bar <-aggregate(df$sales, by=list(segment=df$category), FUN=sum)
col_bar <- c("category","sales")
colnames(df_bar) <-col_bar
df_bar

barplot(df_bar$sales,xlab="category",ylab="sales",col="blue",names.arg=df_bar$category, main="sales by category")
#legend("topright",df_bar$category,fill=c("green","yellow","orange"))

```

if statement

```{r}
if (as.Date("2017-11-17")%in% df$order_date) {
  print("the date is found")
}else{
  print("the date is not found")
}
```

if ....else.. if ...else 
```{r}
if (as.Date("2020-11-01")%in% df$order_date) {
  print("the date is found")
}else if (as.Date("2017-11-17")%in% df$order_date) {
  print ("another date is found !")
}else{
  print("the date is not found!")
}
```

switch case 

```{r}
switch_ex <- switch(
as.numeric(format(as.Date("2017-02-1"), "%m")),
 "January",
 "February",
 "March"
)
print(switch_ex)
 
```
Repeat loop

When the break statement is encountered inside a loop, the loop is immediately
terminated and program control resumes at the next statement following the loop

```{r}

early_date <- as.Date(df$order_date)
print( min(early_date))
x <-as.numeric(format(as.Date(df$order_date),"%y"))
cnt <- min(x)
repeat{
  print(cnt)
cnt <- cnt+1
  if(cnt > max(x)){
    break
  }
}
```

while loop 

```{r}
early_date <- as.Date(df$order_date)
print( min(early_date))
x <-as.numeric(format(as.Date(df$order_date),"%y"))
cnt <- min(x)
while (cnt <max(x)) {
  print(cnt)
  cnt=cnt+1
 
 
}
```

for loop

```{r}
col<- unique(df$segment)
print(col)
for ( i in col){
 print(i)
}

```

The next statement in R programming language is useful when we want to skip
the current iteration of a loop without terminating it.

```{r}
col<- unique(df$category)
print(col)
for ( i in col){
  if(i=="Office Supplies"){
  next
  }
 print(i)
}

```

The dplyr is a powerful R-package to manipulate, clean and summarize unstructured data in a a fast way


```{r}
library(dplyr)

# the distinct function is used to eliminate duplicates based off the order id
df_x <- df%>%distinct(order_id,.keep_all= TRUE)
df_x


```

```{r}

df_select <- distinct(df%>%select(order_id, customer_id,order_date,ship_date,ship_mode,sales,))
#df_select

df_filter <- df_select%>%filter(ship_mode=='First Class') 
#df_filter

df_arrange <- df_filter%>%arrange(desc(ship_date),order_date, )
#head(df_arrange)

df_mutate <- df_arrange%>%mutate(days=as.Date(ship_date)-as.Date(order_date))
#head(df_mutate)

df_group_by <- group_by(df_mutate,days)
df_group_by


```
inner join , right join, full join 

%>% :This operator will forward a value, or the result of an expression, into the next function call/expression

```{r}

df3 <- data.frame(df$order_id,df$ship_mode,df$segment,df$category)
cols3 <-c("order_id","ship_mode","segment","category")
colnames(df3) <-cols3
  
df4 <- data.frame(df$order_id,df$sales,df$quantity,df$discount,df$profit)
cols4 <-c("order_id","sales","quantity","discount","profit")
colnames(df4) <-cols4

inner <- df3 %>% inner_join(df4,by="order_id")
# inner

right <- df3 %>% right_join(df4,by="order_id")
#right

left <-df3 %>% left_join(df4,by="order_id")
#left

outer <- df3 %>% full_join(df4,by="order_id")
outer
 
```